A Motion Capture and Imitation Learning-based Approach to Robot Control

### 2022-06-28 tex-arxiv/article.tex
fixes to proposed approach section; mocap caption

In this paper, we start by introducing some key concepts in imitation learning -- in particular, how they tie in with the autoregressive use of sequence-to-sequence models \ref{sec:prelim}. Then we explore the broad directions research in this field has taken, as well as prior work relating specifically to the two key practical aspects of our work -- the use of motion capture and throwing tasks \ref{sec:related}. This is followed by a detailed explanation of the reasoning behind our approach \ref{sec:approach}, as well as a description of the practical implementation \ref{sec:materials}. Metrics for evaluating performance on the motivating use case are also introduced in subsection \ref{sec:eval}. In the results \ref{sec:results} and discussion \ref{sec:discussion} sections, we compare the performance of the various model types.

@@

Three crucial decisions need to be made when devising an imitation learning-based approach to robot control:
\begin{itemize}
	\item Data collection -- what will serve as the expert policy? How will data be obtained?
	\item Model architecture -- what type of model template will be used to learn the policy? What are its inputs and outputs, what pre- and post-processing steps will these formats call for?
	\item Control method -- how will model outputs be used in robot motion planning or feedback control?
\end{itemize}

@@

Simulation allows recording exact pose information directly at regular time intervals, but setting up the scene necessitates creating a virtual environment where the task can be performed. In addition, for realistic interactions a physics simulation and immersive interface such as VR is required, but even this does not present the human demonstrator with the haptic feedback of actually performing the task in the real world. As such we decided that the benefits of simple setup and inherently natural interaction with the physical scene outweigh the precision and cost advantages of simulated environments, and selected motion capture as the source of demonstration data. For extracting additional information specific to the throwing task, we also track the object thrown and use its pose information to annotate each observation with target coordinates extrapolated from its flight arc and a release timing signal determined by heuristics described in the section on preprocessing \ref{sec:preprocess}.

@@

Outputting pose derivatives (linear and angular velocities) is advantageous as it enables the use of servo controllers, should implementations of such exist, and input-output timing constraints are less stringent than they would be in a strict sequence model that assumes a constant time step. However, a model-in-the-loop control scheme is required~-- forward planning is not possible without a means to integrate model outputs with realistic feedback from the environment. Moreover, for training the model derivatives of the pose need to be estimated from sequential observations, highly susceptible to discretization error.

@@

\begin{figure}
	\centering
	\fbox{\includegraphics[width=16cm]{../img/mocap_process.jpg}}
	\caption{Motion capture equipment and process. }
	\label{fig:fig1}
\end{figure}

>>>>

In this paper, we start by introducing some key concepts in imitation learning -- in particular, how they tie in with the autoregressive use of sequence-to-sequence models (\ref{sec:prelim}). Then we explore the broad directions research in this field has taken, as well as prior work relating specifically to the two key practical aspects of our work -- the use of motion capture and throwing tasks (\ref{sec:related}). This is followed by a detailed explanation of the reasoning behind our approach (\ref{sec:approach}), as well as a description of the practical implementation (\ref{sec:materials}). Metrics for evaluating performance on the motivating use case are also introduced in subsection (\ref{sec:eval}). In the results (\ref{sec:results}) and discussion (\ref{sec:discussion}) sections, we compare the performance of the various model types.

@@

Three crucial decisions need to be made when devising an imitation learning-based approach to robot control:
\begin{itemize}
	\item data collection -- what will serve as the expert policy? How will data be obtained?
	\item Model architecture -- what type of model template will be used to learn the policy? What are its inputs and outputs, what pre- and post-processing steps will these formats call for?
	\item Control method -- how will model outputs be used in robot motion planning or feedback control?
\end{itemize}

@@

Simulation allows recording exact pose information directly at regular time intervals, but setting up the scene necessitates creating a virtual environment where the task can be performed. In addition, for realistic interactions a physics simulation and immersive interface such as VR is required, but even this does not present the human demonstrator with the haptic feedback of actually performing the task in the real world. Finally, whenever data obtained in a simulated environment is used to solve control challenges, the so-called \emph{reality gap} or \emph{sim2real problem} needs to be tackled -- the difference between the responses of simulated and real environments to the same stimuli \cite{hofer2021sim2real}. As such, we decided that the benefits of simple setup and inherently natural interaction with the physical scene outweigh the precision and cost advantages of simulated environments, and selected motion capture as the source of demonstration data. For extracting additional information specific to the throwing task, we also track the object thrown and use its pose information to annotate each observation with target coordinates extrapolated from its flight arc and a release timing signal determined by heuristics described in the section on preprocessing (\ref{sec:preprocess}).

@@

Outputting pose derivatives (linear and angular velocities) is advantageous as it enables the use of servo controllers, and input-output timing constraints are less stringent than they would be in a sample frequency-based model where a constant time step is used to modulate velocities. However, a model-in-the-loop control scheme is required~-- forward planning is not possible without a means to integrate model outputs with realistic feedback from the environment. Moreover, for training the model derivatives of the pose need to be estimated from sequential observations, highly susceptible to discretization error.

@@

\begin{figure}
	\centering
	\fbox{\includegraphics[width=16cm]{../img/mocap_process.jpg}}
	\caption{Motion capture equipment and demonstration acquisition process. }
	\label{fig:fig1}
\end{figure}
